---
title: 'STAT 428 Homework 0: Solution'
author: "Vishesh Gupta"
date: "1/31/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise

Below are a few toy problems. You can type your answers under each problem.
If you like, you can submit your .Rmd and .pdf files on Gradescope. I'll read them.

### Problem 1

(a) Using bullet points, list two reasons you are taking this course.

+ This is one of my major requirements.
+ I want to further understand the importance of statistics and its application in ML while developing my R skills. 

(b) Using a numbered list, tell me two things about you (e.g., what you are interested in, what you do for fun during the pandemic, any questions/concerns about the course). 

(1) I love playing soccer and have my own youtube channel
(2) I have one dog which is a maltese


### Problem 2

What is the PDF of a normal RV $X\sim N(\mu, \sigma^2)$? Write it as a LaTeX equation block.

$$f_X(x) = \frac{1}{\sqrt{2\pi}\sigma}e^{-\frac{(x-\mu)^2}{2\sigma^2}},   ~~~~~~\infty<x<\infty$$


### Problem 3

Make a code chunk: Create a function that takes $x, \mu, \sigma$ as inputs and outputs $f(x)$, the density of $x$ under $N(\mu, \sigma^2)$. Try not to use built-in functions like `dnorm()`.

```{r normal_pdf}

norm_density <- function(x, mu, sigma){
  
  d = 1/(sqrt(2*pi)*sigma)*exp(-(x-mu)^2/(2*sigma^2))
  
  return(d)
}

# Check: does it produce same value as dnorm?

# f(1) for X~N(0,1^2)
norm_density(x=1,mu=0,sigma=1)
dnorm(1,mean = 0,sd = 1)

# f(5) for X~N(3,5^2)
norm_density(x=5,mu=3,sigma=5)
dnorm(5,mean = 3, sd = 5)

```


### Problem 4

Make a code chunk: Using the function from Problem 3, plot the PDF of $X\sim N(0,1)$ between $-4$ and $4$.


```{r sdnorm_plot}

x = seq(-4,4,.01)
fx = norm_density(x,mu=0,sigma=1)
plot(x,fx, type = 'l', xlab = 'x', ylab = 'f(x)')

```




